---
title: "Multilayer analysis"
author:
- Andrzej Jarynowski (ajarynowski@gmail.com)
date: "23 April 2019"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: true
    depth: 3
    fig_caption: yes
    highlight: tango
    number_sections: yes
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multilayer analysis

To introduce the problem we will use the African Swine Fever spread in  Poland case. Each type of relation (pig production supply chain, wild board density, human mobility) defines a network between counties:
_vertices_ (or _nodes_) represent counties 
_edges_ connect adjacent vertices (possibility of transmitting infection.

We use the libraries:
- multinet: an R package to analyze multiplex social networks represented within the more general framework of multilayer networks
- igraph   R package to analyze  social networks represented within the more general framework of disease spread.

Co-founded by OpenMultiMed Cost Action

```{r}
library(multinet)
library(igraph)
```


# Transmission prediction as arrival time differnce

```{r}

load("data.RData")

net=graph_from_adjacency_matrix(matrix,  weighted=TRUE, mode="directed")

#list - ASF introduction times to Polish counties (till 2017) with their characteristics (pig density, forest coverage, coordinates)


V(net)$names=list$county


```


In particular, we explicitly compute a layout that draws counties at geographical cooridinates.

```{r}
V(net)$x=list$x
V(net)$y=list$y
# coordinates (centroids of counties)
#V(net)$District=list$District

lay=matrix(data=c(V(net)$x,V(net)$y), nrow=18, ncol=2)

plot.igraph(net, vertex.size=9, vertex.label=V(net)$names, vertex.label.cex=0.7, edge.arrow.size=0.3, layout=lay )

# first year of outbreak
for (i in 1:length(list[,1])) {
  if (list$arrival_time[i]<365) {
    V(net)[i]$first=1
  }
}
plot.igraph(net, vertex.size=9, vertex.label=V(net)$names, vertex.label.cex=0.7, edge.arrow.size=0.3, vertex.color=c( "pink", "skyblue")[1+(V(net)$first==1)], layout=lay )

```



To compute effective distances, we can use paths detecting algorithms. 
```{r}

# Average path length 
# The mean of the shortest distance between each pair of nodes in the network 
# (in both directions for directed graphs). 
mean_distance(net, directed=F)
mean_distance(net, directed=T)

# We can also find the length of all shortest paths in the graph:
distances(net) # with edge weights
distances(net, weights=NA) # ignore weights

# We can extract the distances to a node or set of nodes we are interested in.
# Here we will get the distance of every counte from the first affected - bialostocki county.
dist.from.bialystok <- distances(net, v=V(net)[names=="bialostocki"], to=V(net), weights=NA)

# Set colors to plot the distances:
oranges <- colorRampPalette(c("dark red", "gold"))
col <- oranges(max(dist.from.bialystok)+1)
col <- col[dist.from.bialystok+1]

plot(net, vertex.color=col, vertex.label=dist.from.bialystok, edge.arrow.size=.6, 
     vertex.label.color="white")

# We can also find the shortest path between specific nodes.
# Say here between bialski county (main regional hub for pig production) with source (first affected county):
infection.path <- shortest_paths(net, 
                            from = V(net)[names=="bialostocki"], 
                            to  = V(net)[names=="bialski"],
                            output = "both") # both path nodes and edges
infection.path

res <- sir(net, beta=5, gamma=1, no.sim=10)
plot(res)

```
# Multi-layer

In the following we will use a network of affected counties (till 2017) with their layers (pig density, forest coverage, human densiry
```{r}
net <-read.ml("poland_18_kopia2.mpx")
net
#write_ml(net,file)

```
Layers by layers alone

```{r out.width = '90%', fig.align='center', fig.align='center', tidy=FALSE}
l <- layout.multiforce.ml(net, w_inter = 0, gravity = 1)
plot(net, 
     layout = l,
     grid = c(1,3),
     vertex.labels.cex = 0.6,
     legend.x="bottomright", legend.inset = c(.05,.05)
     )
```
# Layer-by-layer statistics

A first quantitative comparison of the layers can be done computing basic network measures for each layer.
The _summary_ computes a selection of measures on all the layers, and also on a flattened network.
```{}
summary(net)
```

```{r, echo=FALSE}
kable(summary(net), digits=2, caption = "Network statistics: n: order (number of nodes), m: size (number of edges), n.comp: number of connected components, dens: density, cc: clustering coefficient, apl: average path length, dia: diameter. _flat_ is the combination of all the layers.")
```

Degree distribution

```{r echo=FALSE}
par(mfrow=c(2,2))
hist(degree.ml(net), breaks = max(degree.ml(net)), main="flattened", xlab="degree")
  for (i in 1:num.layers.ml(net)) {
    d <- degree.ml(net, layers = layers.ml(net)[[i]])
    hist(d, breaks = max(d, na.rm = TRUE), main=layers.ml(net)[[i]], col=i, xlab="degree")
}
par(mfrow=c(1,1))
```
The Multinet package provides additional similarity functions.

_layer.comparison.ml_ function that assembles a table with pair-wise comparisons:
```{}
layer.comparison.ml(net, method="jeffrey.degree")
```

```{r, echo=FALSE}
kable(layer.comparison.ml(net, method="jeffrey.degree"), digits=2, caption = "Dissimilarity between degree distributions, computed using the Jeffrey dissimilarity function.")
```
he correlation between the degrees:

```{r}
comp <- layer.comparison.ml(net, method = "pearson.degree")
```

```{r echo = FALSE}
kable(comp, digits=2, caption="Linear correlation between the degree of actors in the two layers, from -1 (top actors in one layer are not active in the other and vice versa) to 1 (top actors in one layer are top actors in the other and vice versa)")
```
Adjacent to the same other actors in different layers:

```{r}
comp <- layer.comparison.ml(net,method="jaccard.edges")
```

```{r echo = FALSE}
kable(comp, digits=2, caption="Overlapping between edges in the two layers. 0: no actors adjacent in one layer are also adjacent in the other. 1: all pairs of actors are either adjacent in both layers or in none.")
```
Which county is the most central actor? Let us see how the most central actors look like when we "unpack" their centrality on the different layers:
```{r echo=FALSE}
kable(
data.frame(wb=degree.ml(net,actors = names(deg),layers = "wb"),
           hum=degree.ml(net,actors = names(deg),layers = "hum"),
           pig=degree.ml(net,actors = names(deg),layers = "pig"),
           flat=deg),caption = "Degree for U4 on each layer",align = "c")
```
#  community detection (multilayer)

Communities could be defined for each layer separetely
```{r tidy=FALSE}
ml_clust <- glouvain.ml(net)

l <- layout.multiforce.ml(net)

plot(net, 
     com = ml_clust, 
     vertex.labels.cex = 0.6,
     layout=l, grid = c(1,3),
     legend.x="bottomright", legend.inset = c(.05, .05)
     )
```
another method clique percolation algorithm:
```{r tidy=FALSE}
ml_clust <- clique.percolation.ml(net, k=5, m=2)

l <- layout.multiforce.ml(net)

plot(net, 
     com = ml_clust, 
     vertex.labels.cex = 0.5,
     layout=l, grid = c(1,3),
     legend.x="bottomright", legend.inset = c(.05, .05)
     )
```
only with communities on at least two layers:
```{r echo = FALSE}
ml_clust <- clique.percolation.ml(net, k=5, m=2)
kable(ml_clust[with(ml_clust, order(cid, actor, layer)), ])

```
Only with communities on at least two layers:
```{r echo = FALSE}

omeg=matrix(0,nrow=3,ncol=3)
omeg[1,1]=0.6
omeg[2,2]=0.5
omeg[3,3]=0.3
omeg[1,2]=0.2
omeg[2,1]=0.2
omeg[2,3]=0.3
omeg[3,2]=0.3
omeg[1,3]=0.1
omeg[3,1]=0.1

ml_clust <- glouvain.ml(net, omega=0.3)
```
Relevance

```{r echo = FALSE}
neighborhood_ml(net, actors = "bialski", layers = "wb")
#  who are not connected to bialski on other layers
xneighborhood_ml(net)
# percentage of neighbors of U3 who are also co-workers
relevance_ml(net)
# redundancy between wb and pig
connective_redundancy_ml(net)
# percentage of neighbors of bialski who would no longer be neighbors by removing this lay

```
evolving according to the external dynamic (importing an edge from the other layer).

```{r}
models_mix <- c(evolution.er.ml(350), evolution.er.ml(200), evolution.er.ml(200))
pr.external <- c(0.2,0.5,0.9)
pr.internal <- c(0.8,.5,0.1)
dependency <- omeg
ml_generated_mix <- grow.ml(380, 650, models_mix, pr.internal, pr.external, dependency)
# multinet::layers_ml(ml_generated_mix)
num.edges.ml(ml_generated_mix, layers1 = "L0")
num.edges.ml(ml_generated_mix, layers1 = "L1")
num.edges.ml(ml_generated_mix, layers1 = "L2")
```

```{r tidy=FALSE}
l <- layout.multiforce.ml(ml_generated_mix, w_inter = 1000, iterations = 200, gravity = .5)

f <- function(x) {neighborhood.ml(ml_generated_mix, actors = x[[1]], layers = x[[2]])}

plot(ml_generated_mix, 
     vertex.labels = "", vertex.cex = apply(vertices.ml(ml_generated_mix),1,f)/3, 
     layout = l 
     )
```



